On the 3-cycle of the logistic map

### 2013-04-18 doc/p3p.tex
improving p3p.tex
improve the tth command

%\section{Onset point}
\textbf{Onset point.}
%
As in Ref. \cite{zhang},
we obtain from $c - b = f(b) - f(a)$ that
$c-b = -(b+a)(b-a)$;
Similar,
$a-c = -(c+b)(c-b)=(c+b)(b+a)(b-a)$.
Now $(b-a) + (c-b) + (a-c) = 0$ means
\[
1 - (b+a) + (c+b)(b+a) = 0.
\]
Cycling variables
$a\rightarrow b$,
$b\rightarrow c$,
$c\rightarrow a$,
and adding up the three distinct versions yields
\begin{equation}
3 - 2X + X^2 + Y = 0.
\label{eq:stair}
\end{equation}
%
On the other hand, the sum of (\ref{eq:abc}) gives
$X = 3R - X_2 = 3R - X^2 + 2Y$ (for $X^2 = X_2 + 2 Y$).
%
Using (\ref{eq:stair}) for $Y$ yields
\begin{equation}
  X^2 - X + 2 - R = 0,
\label{eq:XR}
\end{equation}
This equation has a real root only if $2 - R \le 1/4$.
Thus, $R = 7/4$ (so $r = 1+\sqrt{1+4R} = 1+ \sqrt 8$)
is the onset point of the only real 3-cycle.

@@

The composite map $f(f(f(x)))$ is marginally stable
at the onset (bifurcation) point, so
$\frac{d}{da}f(f(f(a))) = f'(c)f'(b)f'(a) = +1 (-1)$
\cite{saha}.
Since $f'(x) = -2x$, we have
\begin{equation}
  Z = \mp1/8.
\label{eq:Zder}
\end{equation}
%
At the onset point $Z = -1/8$, (\ref{eq:RZ}) becomes
\[
  (R-7/4) (R^2 - R/4 + 7/16) = 0,
\]
whose real solution $R = 7/4$ agrees with the previous solution.
%
At the bifurcation point $Z = 1/8$, we have
\[
  R^3 - 2 R^2 + 9R/8 - 81/64 = 0.
\]
The only real solution is
$
R = \frac{2}{3}
  +\frac{1}{4} \sqrt[3]{ \frac{1915}{54} + \frac{5}{2} \sqrt{201}}
  +\frac{1}{4} \sqrt[3]{ \frac{1915}{54} - \frac{5}{2} \sqrt{201}}.
$

Since $a$, $b$ and $c$ are the distinct roots of
\[
  x^3 - X x^2 + Y x - Z = 0.
\]
The discriminant is
\[
\Delta = X^2 Y^2 - 4 Y^3 - 4 X^3 Z + 18 X Y Z - 27 Z^2
  = (4X^2 - 6X + 9)^2,
\]
where we have used (\ref{eq:Y}) and (\ref{eq:Z}),
and eliminated $R$ by (\ref{eq:XR}).
Thus as long as $X$ is real, $\Delta > 0$, which means there are always three
real roots after the onset $R \ge 7/4$
(Thank Beiye Feng for correspondence on this point).

\begin{thebibliography}{}
\bibitem{zhang} C. Zhang,
  Period three begins,
  \textit{Mathematics Magazine}
  \textbf{83},
  295-297
  (2010).

@@

>>>>

%\section{Onset point}
\textbf{Onset point.}
%
%As in Ref. \cite{zhang},
From (\ref{eq:abc}), we have
$c-b = f(b) - f(a) = -(b+a)(b-a)$;
similar,
$a-c = -(c+b)(c-b)=(c+b)(b+a)(b-a)$.
Now $(b-a) + (c-b) + (a-c) = 0$ means
\[
1 - (b+a) + (c+b)(b+a) = 0.
\]
Cycling variables
$a\rightarrow b$,
$b\rightarrow c$,
$c\rightarrow a$
twice,
and adding up the three versions yields
\begin{equation}
3 - 2X + X^2 + Y = 0.
\label{eq:stair}
\end{equation}
%
On the other hand, the sum of (\ref{eq:abc}) gives
$X = 3R - X_2 = 3R - X^2 + 2Y$ (for $X^2 = X_2 + 2 Y$).
%
Using (\ref{eq:stair}) for $Y$ yields
\begin{equation}
  X^2 - X + 2 - R = 0,
\label{eq:XR}
\end{equation}
This equation has a real root only if $2 - R \le 1/4$.
Thus, $R = 7/4$ (so $r = 1+\sqrt{1+4R} = 1+ \sqrt 8$)
is the onset point of the only real 3-cycle
\cite{saha, bechhoefer, gordon, burm, zhang}.

@@

The composite map $f(f(f(x)))$ is marginally stable
at the onset (bifurcation) point, so
$\frac{d}{da}f(f(f(a))) = f'(c)f'(b)f'(a) = +1 (-1)$
\cite{saha}.
As $f'(x) = -2x$, we get
\begin{equation}
  Z = a b c = \mp1/8.
\label{eq:Zder}
\end{equation}
%
At the onset point $Z = -1/8$, (\ref{eq:RZ}) becomes
\[
  (R-7/4) (R^2 - R/4 + 7/16) = 0,
\]
whose real solution $R = 7/4$ agrees with the previous result.
%
At the bifurcation point $Z = 1/8$, we have
\[
  R^3 - 2 R^2 + 9R/8 - 81/64 = 0.
\]
The only real solution is
$R = \frac{2}{3}
  +\frac{1}{4} \sqrt[3]{ \frac{1915}{54} + \frac{5}{2} \sqrt{201}}
  +\frac{1}{4} \sqrt[3]{ \frac{1915}{54} - \frac{5}{2} \sqrt{201}}$
\cite{gordon, burm}.

Since $a$, $b$ and $c$ are the distinct roots of
\[
  x^3 - X x^2 + Y x - Z = 0.
\]
The discriminant $\Delta$ is a square:
\[
X^2 Y^2 - 4 Y^3 - 4 X^3 Z + 18 X Y Z - 27 Z^2
  = (4X^2 - 6X + 9)^2,
\]
where we have used (\ref{eq:Y}) and (\ref{eq:Z}),
and eliminated $R$ by (\ref{eq:XR}).
Thus, as long as $X$ is real
[which is true after the onset $R \ge 7/4$ by (\ref{eq:XR})],
we have $\Delta > 0$, and the cycle points are real.
(Thank Beiye Feng for suggesting this point).

\begin{thebibliography}{}

@@

\bibitem{bechhoefer} J. Bechhoefer,
  The birth of period 3, revisited,
  \textit{Mathematics Magazine}
  \textbf{69},
  115-118
  (1996).

\bibitem{gordon} W. B. Gordon,
  Period three trajectories of the logistic map,
  \textit{Mathematics Magazine}
  \textbf{69},
  118-120
  (1996).

\bibitem{burm} J. Burm, P. Fishback,
  Period-3 Orbits via Sylvester's Theorem and Resultants,
  \textit{Mathematics Magazine}
  \textbf{74},
  47-51
  (2001).

\bibitem{zhang} C. Zhang,
  Period three begins,
  \textit{Mathematics Magazine}
  \textbf{83},
  295-297
  (2010).
